---
title: "p8105_HW1_tc2899"
output: html_document
author: Thomas Carpino, tc2899
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**HW 1 Instructions:**

**Problem 0.1**

Create a public GitHub repo + local R Project --> complete! 
create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to github_document --> complete! 

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems. --> sure thing!

For this Problem, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible. --> sounds good! 

**Problem 0.2**

This “problem” focuses on correct styling for your solutions to Problems 1 and 2. We will look for:

  - meaningful variable / object names
  - readable code (one command per line; adequate whitespace and indentation; etc)
  - clearly-written text to explain code and results
  - a lack of superfluous code (e.g. no unused variables are defined)
  
**Problem 1:** 

Create a data frame comprised of:

  - A random sample of size 10 from a uniform[0, 5] distribution --> done 
  - A logical vector indicating whether elements of the sample are greater than 2 --> done 
  - A (length-10) character vector --> done 
  - A (length-10) factor vector --> done 
  
```{r question1]}
library(tidyverse)

## set seed to ensure reproducibility

set.seed(98)

#creating data frame

problem1 = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = c(vec_numeric > 2),
  vec_char = c("I", "think", "data", "science", "is", "a", "really", "really", "cool", "class"),
  vec_factor = factor(c("big", "small", "big", "small", "small","big", "small", "big", "small", "small"))
)

mean(problem1$vec_numeric)
mean(problem1$vec_logical)
mean(problem1$vec_char)
mean(problem1$vec_factor)
```

When taking the mean of each variable of the dataframe, vec_numeric and vec_logical both work, whereas vec_char and vec_factor do not work. This is because the mathematical average can be found for the first two variables, where vec_numeric ranges from 0 to 5, and vec_logic ranges from 0 to 1. The other two variables cannot be computed using letters as is - NAs are produced as the result. 

```{r}
as.numeric(problem1$vec_logical)
as.numeric(problem1$vec_char)
as.numeric(problem1$vec_factor)
```

When converting the logical, character, and factor variables ot numeric variables, only logical and factor function properly. The character variable did not convert to numeric - NAs are produced 'by coercion'. 

```{r}
as.numeric(as.factor(problem1$vec_char))
as.numeric(as.character(problem1$vec_factor))
```

When converting logical to character to factor to numeric, the function works properly and provides numeric output. However, when converting factor to character to numeric, the function does not work properly and NAs are produced 'by coercion'. 

**Problem 2:**
  
```{r question2]}
library(tidyverse)

set.seed(98)

problem2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  logical = c(x + y > 0)
  )

problem2b = tibble(
  problem2$x,
  problem2$y,
  problem2$logical,
  numeric = as.numeric(problem2$logical),
  factor = as.factor(problem2$logical)
)
```

Write a short description of your vector using inline R code, including: * the size of the dataset * the mean and median of x * 

The mean of x is `r mean(x)` and its median is `r median(X)`. The proportion of cases for which the logical vector is TRUE is `r mean(problem2$logical)`. 

**Scatterplots / Visualizations**

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

Export your first scatterplot to your project directory using ggsave.
  
