---
title: "p8105_HW1_tc2899"
output: html_document
author: Thomas Carpino, tc2899
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**HW 1 Instructions:**

**Problem 0.1**

Create a public GitHub repo + local R Project

--> complete! 

create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to github_document

--> complete! 

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

--> sure thing!

For this Problem, we will assess adherence to the instructions above regarding repo structure, git commit history, and whether we are able to knit your .Rmd to ensure that your work is reproducible.

--> sounds good! 

**Problem 0.2**

This “problem” focuses on correct styling for your solutions to Problems 1 and 2. We will look for:

  - meaningful variable / object names
  - readable code (one command per line; adequate whitespace and indentation; etc)
  - clearly-written text to explain code and results
  - a lack of superfluous code (e.g. no unused variables are defined)
  
**Problem 1:** 

Create a data frame comprised of:

  - A random sample of size 10 from a uniform[0, 5] distribution --> done 
  - A logical vector indicating whether elements of the sample are greater than 2 --> done 
  - A (length-10) character vector --> done 
  - A (length-10) factor vector --> done 
  
```{r question1]}

library(tidyverse)

## set seed to ensure reproducibility

set.seed(98)

#creating data frame

problem1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = c(vec_numeric > 2),
  vec_char = c("I", "think", "data", "science", "is", "a", "really", "really", "cool", "class"),
  vec_factor = factor(c("big", "small", "big", "small", "small","big", "small", "big", "small", "small"))
)

mean(vec_numeric)
mean(vec_logical)
mean(vec_char)
mean(vec_factor)
```
Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why?

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens? In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?

**Problem 2:**

This problem focuses on plotting and the use of inline R code.

Create a data frame comprised of:
  - x: a random sample of size 1000 from a standard Normal distribution
  - y: a random sample of size 1000 from a standard Normal distribution
  - A logical vector indicating whether the x + y > 0
  - A numeric vector created by coercing the above logical vector
  - A factor vector created by coercing the above logical vector
  
```{r question2]}

## set seed to ensure reproducibility

## the data frame 

set.seed(98)

problem1_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_logical = c(x+y > 0),
  #vec_char = c("I", "think", "data", "science", "is", "a", "really", "really", "cool", "class"),
  #vec_factor = factor(c("big", "small", "big", "small", "small","big", "small", "big", "small", "small"))
)
Write a short description of your vector using inline R code, including: * the size of the dataset * the mean and median of x * the proportion of cases for which the logical vector is TRUE

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

Export your first scatterplot to your project directory using ggsave.
  
