---
title: "p8105_HW1_tc2899"
output: html_document
author: Thomas Carpino, tc2899
---

Here I am loading knitr! 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Problem 1:** 

Create a data frame comprised of:

  - A random sample of size 10 from a uniform[0, 5] distribution --> done 
  - A logical vector indicating whether elements of the sample are greater than 2 --> done 
  - A (length-10) character vector --> done 
  - A (length-10) factor vector --> done 

Here I am loading tidyverse. 

```{r question1]}
library(tidyverse)

## set seed to ensure reproducibility

set.seed(98)

## creating my data frame

problem1 = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = c(vec_numeric > 2),
  vec_char = c("I", "think", "data", "science", "is", "a", "really", "really", "cool", "class"),
  vec_factor = factor(c("big", "small", "big", "small", "small","big", "small", "big", "small", "small"))
)

## here I am taking the mean of each of the different variables. 
mean(problem1$vec_numeric)
mean(problem1$vec_logical)
mean(problem1$vec_char)
mean(problem1$vec_factor)
```

When taking the mean of each variable of the dataframe, vec_numeric and vec_logical both work, whereas vec_char and vec_factor do not work. This is because the mathematical average can be found for the first two variables, where vec_numeric ranges from 0 to 5, and vec_logic ranges from 0 to 1. The other two variables cannot be computed using letters as is - NAs are produced as the result. 

```{r}
as.numeric(problem1$vec_logical)
as.numeric(problem1$vec_char)
as.numeric(problem1$vec_factor)

## Above, I am trying to coerce the different variables (logical, character, and factor) to become numeric variables. 
```

When converting the logical, character, and factor variables ot numeric variables, only logical and factor function properly. The character variable did not convert to numeric - NAs are produced 'by coercion'. 

```{r}
as.numeric(as.factor(problem1$vec_char))
as.numeric(as.character(problem1$vec_factor))

## Above, I am attempting to coerce (char -> factor -> numeric) and (factor -> char -> numeric)
```

When converting logical to character to factor to numeric, the function works properly and provides numeric output. However, when converting factor to character to numeric, the function does not work properly and NAs are produced 'by coercion'. 

**Problem 2:**

```{r question2]}

## Here, I am loading tidyverse once again: 

library(tidyverse)

## Setting the seed:

set.seed(98)

## Creating the dataframe for problem 2:

problem2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  logical = c(x + y > 0),
  numeric = as.numeric(logical),
  factor = as.factor(logical)
  )

```

**Description of the dataframe:**
The size of the dataset is `r dim(problem2)`. The mean of x is `r mean(problem2$x)` and the median of x is `r median(problem2$x)`. The proportion of cases for which the logical vector is TRUE is `r mean(problem2$logical)`. 

**Scatterplots / Visualizations**

Here, I am creating a scatterplot of y vs x (as requested in the homework), and am using the logical vector to determine the color scale. 

```{r xy_scattter1}
ggplot(problem2, aes(x = y, y = x, color = logical)) + geom_point()

ggsave("problem2.pdf")
```

Here is a second scatterplot of y vs x, where the color points are being determined by the numeric vector:

```{r scatterplot2}
ggplot(problem2, aes(x = y, y = x, color = numeric)) + geom_point()
```

Here is a third scatterplot of y vs x, where the color points are being determined by the factor vector:

```{r scaterplot3}
ggplot(problem2, aes(x = x, y = y, color = factor)) + geom_point()
```
  
